---
title: "Housing Price"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### LOAD LIBRARIES AND DATA

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(leaps)
library(margins)
library(earth)
library(ggeffects)
library(vip)

data <- read.csv("housing_5000.csv")

glimpse(data)
```

### A

```{r}
# Identify best subsets
models <- regsubsets(log(PRICEK)~., data = data)
summary(models)

# Select the best subset based on the criterion described
res.sum <- summary(models)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$rss),
  BIC = which.min(res.sum$bic)
)

```

That is best model: log(PRICEK) ~ DSALE + SQFTK + LN_LOTSIZE + CENTRALAIR + GARAGE + FIREPLACE + MICHLAKE + LAKE_RIVER

### B

```{r}
# Descriptive statistics for the best fit model
res.sum$rss[8]
res.sum$adjr2[8]
res.sum$bic[8]

```

### C

```{r}
# Fit the linear regression model
m1 <- lm(log(PRICEK) ~ DSALE + SQFTK + LN_LOTSIZE + CENTRALAIR
         + GARAGE + FIREPLACE + MICHLAKE + LAKE_RIVER, data)
summary(m1)
```

Regression equation:

$log(PRICEK) = 4.419385 - 0.160365*DSALE + 0.189483*SQFTK + 0.122369*LN_LOTSIZE + 0.083921*CENTRALAIR + 0.026255*GARAGE + 0.079700*FIREPLACE + 0.213054*MICHLAKE + 0.123061*LAKE_RIVER$

### D

```{r}
# Adding quadratic effect to SQFT
m2 <- lm(log(PRICEK) ~ DSALE + SQFTK^2 + LN_LOTSIZE + CENTRALAIR
         + GARAGE + FIREPLACE + MICHLAKE + LAKE_RIVER, data)
```


```{r}
summary(margins(m2))
```

### E

```{r, warning=FALSE, message=FALSE}
p <- ggpredict(m1)
plot(p)
```

```{r}
m <- margins_summary(m1)
plot(m)
```

### F

```{r}

# Fit a basic MARS model
mars <- earth(
  log(PRICEK) ~ .,  
  data = data)

# summary of the model
summary(mars)
```

### G

```{r}
# Plot important variables
vip(mars)
```

The identified important variables are not identical to the identified variables previously because MARS find out non-linear relation and linear regression does not do that.








