---
title: "Covid Data Analysis"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data

```{r}
library(tidyverse)
library(caret)
library(ROCR)
library(rpart)
library(rpart.plot)


data <- read.csv("cov.csv")
glimpse(data) # take a look on the data
```

# Initial Questions

Number of people tested positive and negative for covid-19 in this dataset:

```{r}
data %>%
  group_by(corona_result) %>%
  summarise(Freq = n())
```

Number of tested positive = 4553 and negative = 146810.

To remove "other" from corona_result and change positive and negative into 0 and 1 respectively:

```{r}
# Remove other from the result column
data1 <- subset(data, corona_result != "other")

# make positive = 1 and negative = 0

data1 <- mutate(data1, res = recode(corona_result,
                                    "negative" = 0,
                                    "positive" = 1))

# Remove observations of missing values with observations for age_60_and_above and gender

summary(is.na(data1[,8:9]))

```

There are no NA's in these two columns. So missing values can occur with blank spaces. So we shall need to remove those blank spaces to make the data prepared

```{r}

data1 <- subset(data1, age_60_and_above != "" | gender != "")

```


To split the dataset into 90:10 ratio, we can execute the following code:

```{r}
set.seed(111)
library(caTools)
sample <- floor(0.70 * nrow(data1))
train_indx <- sample(seq_len(nrow(data1)), size = sample)
train <- data1[train_indx, ]
test <- data1[-train_indx, ]
```

# Logistic Regression

Fit a logistic model treating the res as dependent variable and the symptoms as independent variable:

```{r}
m1 <- glm(as.factor(res) ~ cough + fever + sore_throat + shortness_of_breath + head_ache + age_60_and_above + gender, family = "binomial", train)
```

Construct a confusion matrix with the threshold value 0.5 in test set.

```{r}
pred_test <- predict(m1, newdata = test, type = "response")
cm <- table(test$res, as.numeric(pred_test >= 0.5))
cm
confusionMatrix(cm)
```


ROC curve of the logistic regression on test set:

```{r}
pred <- prediction(pred_test, test$res)
perf <- performance(pred, "acc")
plot(perf)
auc = performance(pred, measure = "auc")
print(auc@y.values)
```


The coefficients are:

```{r}

coef(m1)

```


```{r}
exp(cbind(Odds_Ratio = coef(m1), confint(m1)))
```


# Decision Tree

Train and generate decision tree using rpart and rpart.plot:

```{r}
m2 <- rpart(res ~ cough + fever + sore_throat + shortness_of_breath + head_ache + age_60_and_above + gender, train, method = "class",
               control = rpart.control(cp = 0.001))
rpart.plot(m2, fallen.leaves = F)
```


```{r}
m2$variable.importance
```


We execute the code to find the answers:

```{r}

pred_test <- (predict(m2, newdata = test, type = "class"))
cm <- table(pred_test, test$res)
confusionMatrix(cm)
```


