---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries and read the data

```{r}
library(tidyverse)
library(caret)
library(CGPfunctions)
library(nnet)

d <- read.csv("ObesityDataSet_raw_and_data_sinthetic.csv")
```

Overview of the data

```{r}
glimpse(d)
```

# Visualization

Impact of gender in obesity.

```{r}
ggplot(d, aes(x = NObeyesdad, y = Gender, fill = Gender)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```


Impact of family history of overweight in obesity.


```{r}
ggplot(d, aes(x = NObeyesdad, y = `family_history_with_overweight`, fill = `family_history_with_overweight`)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Impact of smoking status in obesity.


```{r}
ggplot(d, aes(x = NObeyesdad, y = SMOKE, fill = SMOKE)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```


# Splitting data into train and test sets:

```{r}
set.seed(123)
smp_size <- floor(0.70 * nrow(d))

train_ind <- sample(seq_len(nrow(d)), size = smp_size)

train <- d[train_ind, ]
test <- d[-train_ind, ]
```


# Train a logistic regression model

```{r}
m1 <- multinom(NObeyesdad ~., train)
summary(m1)
```

## Make predictions and find accuracy

```{r}
pred <- predict(m1, type = "class", newdata = test)
head(pred)

cm <- table(pred, test$NObeyesdad)
confusionMatrix(cm)

```

So our model predicts obesity class with 95% accurately.


## Take a look on the important variables

Important variables to determine obesity types.

```{r}
varImp(m1)
```













