---
title: "SVM parameter tuning"
date: "11/30/2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Load libraries

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(kernlab)
library(tidymodels)
library(rsample)
library(doParallel)
library(caret)
```

## Load and slipe the data

```{r}
data <- read.csv("data.csv")

data %>%
  mutate(target = as.factor(target)) -> data
glimpse(data)
data <- rename(data, age = ï..age) # Rename the variabe to avoid garbage characters

```

The following code is to split the dataset into training and test sets

```{r}
set.seed(123) # Set seeds to make it reproducible
split_obj <- initial_split(data = data, prop = 0.7, strata = target)
train <- training(split_obj)
test <- testing(split_obj)
```

## i

Create the recipe

```{r}
rec_obj <-
  recipe(target ~ ., data=train) %>%
  step_impute_median(all_numeric(), -all_outcomes()) %>%
  step_scale(all_numeric(), - all_outcomes()) %>%
  step_unknown(all_nominal(), -all_outcomes()) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  prep()

# Bake

train <- bake(rec_obj, new_data=train)
test <- bake(rec_obj, new_data=test)
```


### Define the tuning specification

```{r}
tuning_specs <- svm_poly(
  cost = tune(), degree = 1
) %>%
  set_mode("classification") %>%
  set_engine("kernlab"); tuning_specs

cost <- expand.grid(cost = c(0.25, 0.5, 0.75, 1, 1.25, 1.5))
cost
```

Resampling with v = 3.

```{r}
cv_folds <- vfold_cv(data = train, v = 3); cv_folds
svm_wf1 <- workflow() %>%
  add_model(tuning_specs) %>% # specifications
  add_formula(target ~ .) # formula
```


### Tuning is faster in parallel

Setting up cores

```{r}
cores1 <- parallel::detectCores(logical = FALSE)
cl <- makeCluster(cores1)
registerDoParallel(cl)
```

### Start Tuning

```{r}
svm_res1 <-  svm_wf1 %>% 
  tune_grid(resamples = cv_folds,
            grid = cost)
```

### Collect the results and optimized model

```{r}
svm_res1 %>% 
  collect_metrics() %>%
  arrange(desc(mean))

final1 <- 
  svm_res1 %>%
  select_best(metric = "accuracy") # best model based on accuracy
final1
```

### Save the final model to an object 

```{r}
model1 <- svm_wf1 %>%
  finalize_workflow(final1) %>%
  fit(data=train) # fit on train data
```

### Prediction and evaluation on training data using model 1

```{r}
p1 <- predict(model1, train, type="prob")
head(p1)
p1 <- p1$.pred_1
pred1 <- p1 >= 0.5
pr1 <- as.numeric(pred1)
confusionMatrix(as.factor(pr1), train$target)

```

### prediction and evaluation on testing data using mdoel1.

```{r}
p2 <- predict(model1, test, type="prob")
head(p2)
p2 <- p2$.pred_1
pred2 <- p2 >= 0.5
pr2 <- as.numeric(pred2)
confusionMatrix(as.factor(pr2), test$target)
```




## iii

Define the tuning specification

```{r}
tuning_specs <- svm_rbf(cost = tune(), rbf_sigma = tune()) %>%
  set_mode("classification") %>%
  set_engine("kernlab"); tuning_specs
```

Define a grid to vary sigma

```{r}
sigma <- expand.grid(rbf_sigma = c(0.1, 1, 2, 3), cost = c(0.25, 0.5, 0.75, 1))
```

Resampling and setting up the workflow

```{r}
cv_folds <- vfold_cv(data = train, v = 3); cv_folds
# Workflow for tuning
svm_wf3 <- workflow() %>%
  add_model(tuning_specs) %>% # specs
  add_formula(target ~ .) # formula
```

### Start Tuning

```{r}
svm_res3 <-  svm_wf3 %>% 
  tune_grid(resamples = cv_folds,
            grid = sigma)
```


### Collect the results and optimized model

```{r}
svm_res3 %>% 
  collect_metrics() %>%
  arrange(desc(mean))

final3 <- 
  svm_res3 %>%
  select_best(metric = "accuracy") # best model based on accuracy
final3
```

### Save the final model to an object 

```{r}
model3 <- svm_wf3 %>%
  finalize_workflow(final3) %>%
  fit(data=train) # fit on train data
```

### Prediction and evaluation on training data using model 3

```{r}
p5 <- predict(model3, train, type="prob")
head(p5)
p5 <- p5$.pred_1
pred5 <- p5 >= 0.5
pr5 <- as.numeric(pred5)
confusionMatrix(as.factor(pr5), train$target)

```

### Prediction and evaluation on training data using model 2

```{r}
p6 <- predict(model3, test, type="prob")
head(p6)
p6 <- p6$.pred_1
pred6 <- p6 >= 0.5
pr6 <- as.numeric(pred6)
confusionMatrix(as.factor(pr6), test$target)
```









